\section{Modeling}
\label{sec:modeling}
We incorporate the modeling of the original paper, which is for wildfire,
into our context of high-building fire. 
We extend the modeling to include smoke at the spreading layer, as well as 
another allocation aspect of rescue teams. 
Despite our effort, more advanced modeling incorporating more factors is still needed
to better simulate the real-world scenario.

\subsection{Model Assumptions}
We approach this spatial allocation problem as a constrained optimization problem, 
following the modeling of the original paper. 
We construct our overall cost function with three separate objective functions: 
the total population lost, the total property loss, and the total cost of unit deployment,
in order of importance. 
To reflect this, we adopt the lexicographic (sequential) optimization strategy, 
which minimizes the objective functions in order. 

We model the fire and smoke spread as in the original paper,
while suppression is a process during which the availability of units differs. 
The spread model estimates the time at which the fire and smoke will reach each cell, 
together with the population and property value in each cell, are given to the solver to 
find the optimal solution. 

We build the solver on the assumption that the optimal allocation is guided by the on-ground
development, including the cell conditions, population in each cell, 
property value of each cell, and the accessibility of each cell. 
The cell condition is determined by the fire and smoke spread arrival time and 
allocation decision in previous time-step, which results an update of situation and
changes in these conditions itself. 

\subsection{Model Formulation}
The model includes the following sets, parameters, decision variables, objective functions,
and constraints described below

\subsubsection{Set}
We construct the following sets in our model: 

\begin{itemize}
    \item $A$: the set of cells, indexed by $a$ 
    \item $I$: the set of resources type, indexed by $i$. 
    \item $T$: the set of time steps, indexed by $t$.
\end{itemize}

\subsubsection{Parameters}
We define the following parameters for constraint in our model: 

\begin{itemize}
    \item $B$: the total budget for unit deployment.

    \item $P_a$: the population in cell $a$ 
    \item $J_a$: the minimum suppression rate in cell $a$. 
    \item $V_a$: the property value in cell $a$.

    \item $G_a$: the accessibility of cell $a$. It is either -1, 0, 1, or 2, meaning inaccessible, land only, air only, or both, respectively. 
    \item $G_{i,a}$: the binary parameter defining whether cell $a$ is accessible to unit $i$. 

    \item $C_i$: the cost of deploying one unit of resource type $i$. 
    \item $S_i$: the suppression rate of resource type $i$.
    \item $R_i$: the rescue rate of resource type $i$.

    \item $K_{i,t}$: the availability of resource type $i$ at time step $t$.
    
    \item $H_{a,t}$: the binary parameter defining whether cell $a$ will be high-risk at fire at time step $t$. 
    \item $F_{a,t}$: the binary parameter defining whether cell $a$ will be high-risk at smoke at time step $t$. 
\end{itemize}

\subsubsection{Decision Variables}
We define the following decision variables in our model:

\begin{itemize}
    \item $x_{i,a,t}$: the number of unit $i$ deployed to cell $a$ at time step $t$.
    
    \item $\mu_{a,t}$: the binary variable defining whether the fire in cell $a$ is high-risk at time step $t$.
    \item $k_{a,t}$: the binary variable defining whether the fire in cell $a$ is untreatable at time step $t$.
    \item $\phi_{a,t}$: the binary variable defining whether the fire in cell $a$ is treated at time step $t$.

    \item $\sigma_{a,t}$: the binary variable defining the smoke in cell $a$ is high-risk at time step $t$.
    \item $f_{a,t}$: the binary variable defining whether the population in cell $a$ is untreatable at time step $t$.
    \item $\psi_{a,t}$: the binary variable defining whether the population in cell $a$ is rescued at time step $t$.
    
    \item $\rho_{a,t}$: the binary variable defining
    \item $z_{i,a,t}$: the binary variable defining whether resource $i$ is deployed to cell $a$ at time step $t$.
\end{itemize}

\subsubsection{Objective Functions}
We define the following objective functions for our model:

\begin{itemize}
    \item $O_1$: the total population at risk
    $$O_1 = \sum_{a \in A} (P_a \sum_{t \in T} \rho_{a,t} ) $$

    \item $O_2$: the total property value lost
    $$O_2 = \sum_{a \in A} (V_a \sum_{t \in T} z_{a,t} ) $$

    \item $O_3$: the total cost of all unit deployment
    $$O_3 = \sum_{i \in I} \sum_{a \in A} \sum_{t \in T} x_{i,a,t} C_i $$
\end{itemize}

The objective functions are optimized in subjection to the constraints. 

\subsubsection{Constraints}
We define the following constraints that express the general conditions and assumptions
and bound the allocation of resources in our model: 

\begin{itemize}
    \item The total cost of unit deployment must not exceed the budget 
    $$\sum_{i \in I} \sum_{a \in A} \sum_{t \in T} x_{i,a,t} C_i \leq B $$

    \item The number of units of each resource type deployed to each cell at each time step 
    must not exceed the availability of that resource type at that time step
    $$\sum_{a \in A} x_{i,a,t} \leq K_{i,t}, \quad \forall i \in I, t \in T$$

    \item The total suppression rate of all resources sent to each cell
    must be sufficient to contain the fire in that cell. 
    $$\sum_{i \in I} \sum_{t \in T} x_{i,a,t} S_i \geq \phi_{a,t} J_a, \quad \forall a \in A, t \in T$$

    \item The total rescue rate of all resources sent to each cell
    must be sufficient to rescue the whole population in that cell. 
    $$\sum_{i \in I} \sum_{t \in T} x_{i,a,t} R_i \geq \psi_{a,t} P_a, \quad \forall a \in A, t \in T$$

    \item Units are only deployed to a cell if that cell is accessible for these units. 
    $$z_{i,a,t} \leq G_{i,a} $$
    $$x_{i,a,t} \leq max_{i}K_{i,t} z_{i,a,t}$$

    \item Only high-risk cells are treated with resources. 
    $$\phi_{a,t} \leq \mu_{a,t}, \quad \forall a \in A, \forall t \in T$$
    $$\psi_{a,t} \leq \sigma_{a,t}, \quad \forall a \in A, \forall t \in T$$

    \item High-risk cells in this time step will become untreatable cells in the next time step
    if they are not treated with resources.
    $$k_{a,t} \leq \mu_{a,t-1} - \phi_{a,t-1}, \quad \forall a \in A, \forall t \in (1,T)$$
    $$f_{a,t} \leq \sigma_{a,t-1} - \psi_{a,t-1}, \quad \forall a \in A, \forall t \in (1,T)$$

    \item Each cell can only a high-risk cell, an untreatable cell, or a treated cell at most once.
    $$\sum_{t \in T} \mu_{a,t} \leq 1, \quad \forall a \in A, \forall t \in T$$ 
    $$\sum_{t \in T} \sigma_{a,t} \leq 1, \quad \forall a \in A, \forall t \in T$$
    $$\sum_{t \in T} k_{a,t} \leq 1, \quad \forall a \in A, \forall t \in T$$ 

    $$\sum_{t \in T} f_{a,t} \leq 1, \quad \forall a \in A, \forall t \in T$$ 
    $$\sum_{t \in T} \phi_{a,t} \leq 1, \quad \forall a \in A, \forall t \in T$$ 
    $$\sum_{t \in T} \psi_{a,t} \leq 1, \quad \forall a \in A, \forall t \in T$$ 

    \item Treated or untreatable cells must previously be high-risk cells.
    $$\sum_{t \in T} k_{a,t} \leq \sum_{t \in T} \mu_{a,t}, \forall a \in A, t \in T$$
    $$\sum_{t \in T} \phi_{a,t} \leq \sum_{t \in T} \mu_{a,t}, \forall a \in A, t \in T$$

    $$\sum_{t \in T} f_{a,t} \leq \sum_{t \in T} \sigma_{a,t}, \forall a \in A, t \in T$$
    $$\sum_{t \in T} \psi_{a,t} \leq \sum_{t \in T} \sigma_{a,t}, \forall a \in A, t \in T$$

    \item A high-risk cell can only be either treated or untreatable in the end, but not both.
    $$\sum_{t \in T} k_{a,t} + \sum_{t \in T} \phi_{a,t} \leq 1, \forall a \in A, t \in T$$
    $$\sum_{t \in T} f_{a,t} + \sum_{t \in T} \psi_{a,t} \leq 1, \forall a \in A, t \in T$$   
    
    \item Each high-risk cell must be treated or untreatable in the end. 
    $$\sum_{t \in T} k_{a,t} + \sum_{t \in T} \phi_{a,t} \geq \sum_{t \in T} \mu_{a,t}, \forall t \in T$$
    $$\sum_{t \in T} f_{a,t} + \sum_{t \in T} \psi_{a,t} \geq \sum_{t \in T} \sigma_{a,t}, \forall t \in T$$

    \item Fire and smoke spread must be consistent with the spread model.  
    $$10 \mu_{a,t} = H_{a,t} \sum_{a' \in neighbors(a)} k_{a',t}$$
    $$10 (1-\mu_{a,t}) \geq 1 - (H_{a,t} \sum_{a' \in neighbors(a)} k_{a',t})$$

    $$10 \sigma_{a,t} = F_{a,t} \sum_{a' \in neighbors(a)} f_{a',t}$$
    $$10 (1-\sigma_{a,t}) \geq 1 - (F_{a,t} \sum_{a' \in neighbors(a)} f_{a',t})$$

    $$\quad \forall a \in A, \forall t \in (1, T)$$

    \item People at risk
    $$\rho_{a,t} \geq \mu_{a,t} - \phi_{a,t}$$
    $$\rho_{a,t} \geq \sigma_{a,t} - \psi_{a,t}$$

\end{itemize}