\section{Related Work}
\label{sec:related-work}

\subsection{Resource Allocation in Fire Departments (RAFD)}
In urban contexts, Resource Allocation in Fire Departments (RAFD) is recognized as a strategic level of public sector management that aims to minimize social costs, such as loss of life and property. 
Structured literature reviews of RAFD demonstrate that mathematical optimization, specifically Integer Programming (IP) and Data Envelopment Analysis (DEA), are the most frequently applied approaches. 
However, the vast majority of urban fire optimization models focus on macro-level coverage problems. 
These studies primarily address the optimal spatial location of fire stations and the dispatch routing of fleets to satisfy coverage demands across jurisdictions, paying special attention to Highly Vulnerable Areas (HVAs) such as hospitals and schools. 
Furthermore, current RAFD studies predominantly utilize historical incident data or static risk scores to predict future demands, rather than adapting to real-time, evolving incident dynamics. 
Consequently, there is a distinct lack of optimization models designed for the active, micro-level tactical deployment of personnel inside a building during an active fire.

\subsection{Wildfire Suppression Optimization Models}
Conversely, operations research in wildland fire management has successfully developed highly dynamic optimization algorithms that adapt to real-time hazard propagation. 
Models frequently utilize Mixed-Integer Linear Programming (MILP) to optimize the allocation of limited ground resources and brigades during the suppression stage. 
The MORA framework is a leading DSS in this domain, formulating resource deployment as a Constrained Optimization Problem (COP). 
Unlike weighted-sum models, MORA employs a lexicographic (sequential) optimization strategy: it strictly prioritizes minimizing the population exposed to fire risk, followed by minimizing property and land losses, and finally minimizing the operational cost of the deployed resources. 
Despite their advanced dynamic routing capabilities and strict operational constraints, models like MORA are inherently designed for 2D geographical landscapes divided into raster cells. 
They have not yet been adapted to the 3D node-based topology of modern high-rises or the specific indoor hazards associated with urban structural fires.

\subsection{IoT, BIM, and Digital Twins in Building Fire Safety}
Recent advancements in smart building design have revolutionized fire risk monitoring through the integration of the Internet of Things (IoT) and Building Information Modeling (BIM). 
Modern frameworks deploy networks of IoT sensors---capturing real-time data on temperature, smoke density, and carbon monoxide (CO) concentrations---which are then visualized within a 3D BIM environment. 
These technologies have proven highly effective for early fire detection, allowing for the calculation of dynamic metrics like the Improved Risk Index (IRI) based on architectural design, occupant density, and real-time environmental conditions. 
The concept of the ``Digital Twin'' is also emerging in fire safety engineering, offering real-time virtual replicas of building conditions to aid emergency management. 

However, significant deficiencies persist. 
While these smart building frameworks excel at passive data collection, hazard visualization, and theoretical evacuation modeling, they lack active, algorithmic decision-making capabilities. 
They provide sensory input but do not possess the integrated mathematical solvers required to actively compute the optimal tactical deployment of rescue and suppression squads.

\subsection{Evacuation Dynamics and the ``Smoke Factor''}
During an indoor fire, the rapid propagation of toxic smoke and gases presents the most immediate threat to both occupants and first responders. 
A recent scoping review on human behavior in fires identified significant research gaps regarding how environmental factors---specifically irritant and toxic smoke---affect movement speed, wayfinding, and decision-making during building evacuations. 
Current mathematical models often approximate evacuation solutions without adequately accounting for the dynamic expansion of high-risk smoke zones, leaving an unsafe discrepancy between theoretical plans and actual emergency outcomes. 

To ensure responder safety and evacuation success, it is critical that routing algorithms proactively avoid hazardous, smoke-filled environments. 
By introducing a dynamic ``smoke factor'' as a strict constraint within an optimization solver, response models can realistically capture the hazards of indoor rescue planning, deliberately preventing responder engagement in compromised or unsurvivable locations. 
MORAv2 builds upon these identified gaps by synthesizing the mathematical rigor of wildfire MILP solvers with the real-time hazard awareness of indoor sensor networks.